---
title: "SCDL3992: Independent Simulations"
date: today
author: "Calvin L."
format:
  html:
    theme: litera
    embed-resources: true 
    code-fold: true
    code-tools: true
    toc: true
    toc-location: right
    number-sections: true
    link-external-newwindow: true
    link-external-icon: true
    math: katex
---

```{r open, warning = F, message = F}
knitr::opts_chunk$set(warning = F, message = F, fig.align='center', fig.dim=c(10, 6))

# Libraries
library(qvalue)
library(tidyverse)
library(knitr)
library(patchwork)
library(mvtnorm)
library(glue)
library(future)
library(furrr)

# Functions
source("simulate.R")
source("metrics.R")
source("FDR_bound_new.R")
```

# Average FDR vs FDP

Take $M=1000$ iterations, $n=1000$ sample size, $\pi_0=0.8$, $\alpha=\{0.01, 0.05, 0.1, 0.2\}$ for now, $\tau=0.25$ and $c=1$.

```{r FDR_FDP-gen}
# Parameters
B = 1000
n = 1000
pi0 = 0.8
params = c(0.8, 0, 1, 0.4, 4)
alpha = c(0.01, 0.05, 0.1, 0.2)
tau = 0.25
c = 1

outfile = "rds/FDP_FDR1.rds"

if (file.exists(outfile)) {
  obj = readRDS(outfile)
  FDR = obj$FDR
  FDP = obj$FDP
} else {
  # pre-allocate
  FDR = matrix(NA_real_, nrow = B, ncol = length(alpha))
  FDP = matrix(NA_real_, nrow = B, ncol = length(alpha))
  colnames(FDR) = as.character(alpha)
  colnames(FDP) = as.character(alpha)

  # simulate
  for (i in seq_len(B)) {
    sim   = generate_p(n = n, p = pi0, seed = i,
                       u_min = 0, u_max = 1, beta_a = 0.4, beta_b = 4)
    p     = sim$p
    nulls = sim$is_null

    for (j in seq_along(alpha)) {
      res = metrics_FDR_bound(p, alpha[j], tau, nulls, c = c)
      FDP[i, j] = res$FDP
      FDR[i, j] = res$FDR_bound_raw   # raw bound-estimator at cutoff
    }
  }

  saveRDS(
    list(
      FDR = FDR,
      FDP = FDP,
      meta = list(B = B, n = n, pi0 = pi0, alpha = alpha, tau = tau, c = c)
    ),
    outfile
  )
}
```

We want to check if the average FDR bound is greater than the average FDP.

```{r table-averages}
# Take col means
FDR_means = colMeans(FDR)
FDP_means = colMeans(FDP)
colnames(FDP) = colnames(FDR) = sprintf("%.2f", alpha)

# Display table
df = data.frame("FDR" = FDR_means, "FDP" = FDP_means) %>% 
  mutate(FDR = round(FDR, 5), FDP = round(FDP, 5)) %>% rename("FDR Bound" = "FDR")

kable(df, align = "ccc", caption = "Average FDR Bound vs Average FDP")
```

## Box Plot of Differences

```{r box-diff}
diff = FDR - FDP
params_txt = paste0("(", paste(params, collapse = ", "), ")")

alpha_labels = c("0.01", "0.05", "0.1", "0.2")
stopifnot(ncol(diff) == length(alpha_labels))

df_diff = data.frame(
  value = as.vector(diff),
  alpha = factor(rep(alpha_labels, each = nrow(diff)), levels = alpha_labels)
)

p_diff = ggplot(df_diff, aes(x = "", y = value)) +
  geom_boxplot() +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.6, color = "red") +
  facet_wrap(~ alpha, nrow = 1) +
  labs(
    title = "Difference between Average FDR_bound and FDP",
    subtitle = glue("parameters = {params_txt}, n = {n}, tau = {tau}, c = {c}, M = {B}"),
    x = NULL,
    y = "Difference (FDR_bound - FDP)"
  ) +
  theme_linedraw(base_size = 13)

p_diff
```

## Plot of FDR Bound against FDP

Line plots of FDR Bound (raw) vs FDP.

```{r alpha-gen}
alphas = seq(0.01, 0.99, by = 0.01)
iters = 100

outfile = "rds/FDP_FDR2.rds"

if (file.exists(outfile)) {
  obj = readRDS(outfile)
  FDR = obj$FDR
  FDP = obj$FDP
} else {
  # Pre-allocate
  FDR = matrix(NA_real_, nrow = iters, ncol = length(alphas))
  FDP = matrix(NA_real_, nrow = iters, ncol = length(alphas))
  colnames(FDR) = as.character(alphas)
  colnames(FDP) = as.character(alphas)

  for (i in seq_len(iters)) {
    sim   = generate_p(n = n, p = pi0, seed = i)
    p     = sim$p
    nulls = sim$is_null

    tmp = lapply(alphas, function(a) metrics_FDR_bound(p, a, tau, nulls, c = c))
    FDP[i, ] = vapply(tmp, function(x) x$FDP,     numeric(1))
    FDR[i, ] = vapply(tmp, function(x) x$FDR_bound_raw, numeric(1))
  }

  saveRDS(
    list(
      FDR = FDR,
      FDP = FDP,
      meta = list(B = B, n = n, pi0 = pi0, tau = tau, c = c, alphas = alphas)
    ),
    outfile
  )
}
```

```{r FDR_FDP-plot}
df = data.frame(alpha = alphas, FDP_mean = colMeans(FDP, na.rm = TRUE),
                FDR_bound_mean = colMeans(FDR, na.rm = TRUE))

df = subset(df, alpha <= 0.25)

ggplot(df, aes(x = alpha)) +
  geom_line(aes(y = FDR_bound_mean, linetype = "FDR_bound", color = "FDR_bound"), linewidth = 1) +
  geom_line(aes(y = FDP_mean,        linetype = "FDP",        color = "FDP"),        linewidth = 1) +
  scale_linetype_discrete(name = "Metric") +
  scale_color_manual(
    name = "Metric",
    values = c("FDR_bound" = "#009E73",  # green
               "FDP"        = "#D55E00") # orange
  ) +
  labs(x = expression(alpha), y = "Average FDR Bound / FDP",
       title = glue("FDR bound vs FDP across α"), 
       subtitle = glue("parameters = {params_txt}, n = {n}, tau = {tau}, c = {c}, M = {B}")) +
  theme_linedraw() +
  guides(linetype = "none", color = guide_legend(title = "Metric"))
```

# Simulations

The goal is to simulate and test different levels of $\tau, c, n, B$ as below and conclude with a plot of Power and number of discoveries (against BH, Storey and BY). When testing a parameter, the rest use the default setting, for a reasonable run-time I had to keep default iterations to $B=100$. 

| Parameter/Test        | Default     | Sim Values        |
| --------------------- | ----------- | ----------------- |
| $\tau$                | 0.25        | 0.1, 0.175, 0.25  |
| $c$                   | 1           | 0, 0.5, 1, 2      |
| Sample Size $n$       | 1000        | 1000, 5000, 10000 |
| Iterations            | 100         | 100, 500, 1000    |
| FDR                   | (use above) |                   |
| Power                 | (use above) |                   |
| Number of discoveries | (use above) |                   |

The data generation parameters stay the same as $(\pi_0, u_1, u_2, a, b)=(0.8, 0, 1, 0.4, 4)$ and $\alpha$ goes from $0.01\to0.99$ over each simulation.

To help save run time and see the trade-offs and differences between different values of a particular parameter, I will plot the three different parameter values on the same plot but just for `FDR_bound`, then later on I will include the last three with reference to BH, BY, and Storey.

```{r default-settings}
# Default parameters
n = 1000
params = c(0.8, 0, 1, 0.4, 4)
tau = 0.25
c = 1
alphas  = seq(0.01, 0.99, by = 0.01)
iters = 100
seed0 = 1L
```

## $\tau$ Simulation

```{r tau-results, warning = F, message = F}
taus = c(0.1, 0.175, 0.25)

# Use all but one core
future::plan(multisession, workers = max(1, parallel::detectCores() - 1))
outfile = "rds/tau_sim.rds"

if (file.exists(outfile)) {
  res_alpha = readRDS(outfile)
} else {
  res_alpha = future_map2_dfr(
    alphas, seq_along(alphas),
    ~{
      map_dfr(taus, function(tt) {
        out = simulate(
          n, alpha = .x, tau = tt, params = params, B = iters,
          seed.offset = seed0 + .y, structure = "ind"
        )
        dplyr::mutate(out, alpha = .x, tau = tt)
      })
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
  saveRDS(res_alpha, outfile)
}

# Keep only FDR_bound + compute MC SEs (uses iters replicates)
df_tau = res_alpha %>%
  dplyr::filter(Method == "FDR_bound") %>%
  dplyr::mutate(
    FDP_mcse   = FDP_sd   / sqrt(iters),
    Power_mcse = Power_sd / sqrt(iters)
  ) %>%   mutate(
    tau_f = factor(
      tau,
      levels = sort(unique(tau)),
      labels = paste0("τ = ", sort(unique(tau)))
    )
  )

```

```{r tau-plot}
df_tau = subset(df_tau, alpha <= 0.25)
# Color palette mapped to the τ factor levels (color-blind friendly)
tau_levels = levels(df_tau$tau_f)
base_cols = c("#0072B2", "#009E73", "#D55E00")  # blue/green/orange
cols_tau = setNames(base_cols[seq_along(tau_levels)], tau_levels)

params_txt = paste0("(", paste(params, collapse = ", "), ")")

fdr_tau = ggplot(df_tau, aes(x = alpha, y = FDP_mean, color = tau_f, fill = tau_f)) +
  geom_ribbon(
    aes(ymin = pmax(0, FDP_mean - FDP_mcse),
        ymax = pmin(1, FDP_mean + FDP_mcse)),
    alpha = 0.12, color = NA
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 1) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "τ", values = cols_tau) +
  scale_fill_manual(name = "τ", values = cols_tau) +
  labs(
    title = glue("Independent p-values | τ Simulations"),
    x = expression(alpha),
    y = "Empirical FDR",
    color = "τ", fill = "τ",
    subtitle = "FDR vs α",
  ) +
  theme_linedraw(base_size = 13) + theme(legend.position = "none")

# Power

power_tau = ggplot(df_tau, aes(x = alpha, y = Power_mean, color = tau_f, fill = tau_f)) +
  geom_ribbon(
    aes(ymin = pmax(0, Power_mean - Power_mcse),
        ymax = pmin(1, Power_mean + Power_mcse)),
    alpha = 0.12, color = NA
  ) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "τ", values = cols_tau) +
  scale_fill_manual(name = "τ", values = cols_tau) +
  labs(
    title = NULL,
    x = expression(alpha),
    y = "Power",
    color = "τ", fill = "τ",
    subtitle = "Power vs α"
  ) +
  theme_linedraw(base_size = 13)

tau_sim = fdr_tau + power_tau
```

## $c$ Simulation

```{r c-results}
tau = 0.25 # default
cs = c(0, 0.5, 1, 2)

outfile = "rds/c_sim.rds"
future::plan(multisession, workers = max(1, parallel::detectCores() - 1))

if (file.exists(outfile)) {
  res_alpha = readRDS(outfile)
} else {
  res_alpha = furrr::future_map2_dfr(
    alphas, seq_along(alphas),
    ~{
      purrr::map_dfr(cs, function(cc) {
        out = simulate(
          n, alpha = .x, tau = tau, params = params, B = iters,
          c = cc, seed.offset = seed0 + .y, structure = "ind"
        )
        dplyr::mutate(out, alpha = .x, c = cc)
      })
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
  saveRDS(res_alpha, outfile)
}

# tidy
df = res_alpha %>%
  dplyr::filter(Method == "FDR_bound") %>%
  dplyr::mutate(
    FDP_mcse   = FDP_sd   / sqrt(iters),
    Power_mcse = Power_sd / sqrt(iters)
  )
```

```{r c-plot}
df_c = res_alpha %>%
  filter(Method == "FDR_bound") %>%
  mutate(
    FDP_mcse   = FDP_sd   / sqrt(iters),
    Power_mcse = Power_sd / sqrt(iters),
    c_f = factor(
      c,
      levels = sort(unique(c)),
      labels = paste0("c = ", sort(unique(c)))
    )
  )

df_c_zoomed = subset(df_c, alpha <= 0.25)

# palette for c levels
c_levels = levels(df_c_zoomed$c_f)
base_cols = c("#0072B2", "#009E73", "#D55E00", "slateblue")
cols_c = setNames(base_cols[seq_along(c_levels)], c_levels)

params_txt = paste0("(", paste(params, collapse = ", "), ")")

fdr_c = ggplot(df_c_zoomed, aes(x = alpha, y = FDP_mean, color = c_f, fill = c_f, group = c_f)) +
  geom_ribbon(
    aes(ymin = pmax(0, FDP_mean - FDP_mcse),
        ymax = pmin(1, FDP_mean + FDP_mcse)),
    alpha = 0.12, color = NA
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 1) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "c", values = cols_c) +
  scale_fill_manual(name = "c", values = cols_c) +
  labs(
    title = glue("Independent p-values | c Simulations"),
    x = expression(alpha),
    y = "Empirical FDR",
    color = "c", fill = "c",
    subtitle = "FDR vs α"
  ) +
  theme_linedraw(base_size = 13) + theme(legend.position = "none")

# Power

power_c = ggplot(df_c_zoomed, aes(x = alpha, y = Power_mean, color = c_f, fill = c_f, group = c_f)) +
  geom_ribbon(
    aes(ymin = pmax(0, Power_mean - Power_mcse),
        ymax = pmin(1, Power_mean + Power_mcse)),
    alpha = 0.12, color = NA
  ) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "c", values = cols_c) +
  scale_fill_manual(name = "c", values = cols_c) +
  labs(
    title = NULL,
    x = expression(alpha),
    y = "Power",
    color = "c", fill = "c",
    subtitle = "Power vs α"
  ) +
  theme_linedraw(base_size = 13)

c_sim = fdr_c + power_c
```

## $n$ Simulation

```{r n-results}
c = 1 # default
ns = c(1000, 5000, 10000)

outfile = "rds/n_sim.rds"

future::plan(multisession, workers = max(1, parallel::detectCores() - 1))

if (file.exists(outfile)) {
  res_alpha = readRDS(outfile)
} else {
  res_alpha = furrr::future_map2_dfr(
    alphas, seq_along(alphas),
    ~{
      purrr::map_dfr(ns, function(nn) {
        out = simulate(
          n = nn, alpha = .x, tau = tau, params = params, B = iters,
          c = c, seed.offset = seed0 + .y, structure = "ind"
        )
        dplyr::mutate(out, alpha = .x, n = nn)
      })
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
  saveRDS(res_alpha, outfile)
}

# tidy
df = res_alpha %>%
  dplyr::filter(Method == "FDR_bound") %>%
  dplyr::mutate(
    FDP_mcse   = FDP_sd   / sqrt(iters),
    Power_mcse = Power_sd / sqrt(iters)
  )
```

```{r n-plot}
# prep: keep only FDR_bound rows & add factor for n
df_n = res_alpha %>%
  filter(Method == "FDR_bound") %>%
  mutate(
    FDP_mcse   = FDP_sd   / sqrt(iters),
    Power_mcse = Power_sd / sqrt(iters),
    n_f = factor(
      n,
      levels = sort(unique(n)),
      labels = paste0("n = ", sort(unique(n)))
    )
  )

df_n_zoomed = subset(df_n, alpha <= 0.25)

# 3-color palette for the three n levels
n_levels = levels(df_n_zoomed$n_f)
base_cols = c("#0072B2", "#009E73", "#D55E00")  # blue / green / orange
cols_n = setNames(base_cols[seq_along(n_levels)], n_levels)

params_txt = paste0("(", paste(params, collapse = ", "), ")")

fdr_n = ggplot(df_n_zoomed, aes(x = alpha, y = FDP_mean, color = n_f, fill = n_f, group = n_f)) +
  geom_ribbon(
    aes(ymin = pmax(0, FDP_mean - FDP_mcse),
        ymax = pmin(1, FDP_mean + FDP_mcse)),
    alpha = 0.12, color = NA
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 1) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "n", values = cols_n) +
  scale_fill_manual(name = "n", values = cols_n) +
  labs(
    title = "Independent p-values | n Simulations",
    x = expression(alpha),
    y = "Empirical FDR",
    color = "n", fill = "n",
    subtitle = "FDR vs α"
  ) +
  theme_linedraw(base_size = 13) + theme(legend.position = "none")

power_n = ggplot(df_n_zoomed, aes(x = alpha, y = Power_mean, color = n_f, fill = n_f, group = n_f)) +
  geom_ribbon(
    aes(ymin = pmax(0, Power_mean - Power_mcse),
        ymax = pmin(1, Power_mean + Power_mcse)),
    alpha = 0.12, color = NA
  ) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "n", values = cols_n) +
  scale_fill_manual(name = "n", values = cols_n) +
  labs(
    title = NULL,
    x = expression(alpha),
    y = "Power",
    color = "n", fill = "n",
    subtitle = "Power vs α"
  ) +
  theme_linedraw(base_size = 13)

n_sim = fdr_n + power_n
n_sim
```

## $M$ Simulation

Using a different seed from before, now the plot for $M$ looks like this.

```{r B-results}
n = 1000 # default
Bs = c(100, 500, 1000)
seed0 = 188

outfile = "rds/B_sim2.rds"

future::plan(multisession, workers = max(1, parallel::detectCores() - 1))

if (file.exists(outfile)) {
  res_alpha = readRDS(outfile)
} else {
  res_alpha = furrr::future_map2_dfr(
    alphas, seq_along(alphas),
    ~{
      purrr::map_dfr(Bs, function(BB) {
        out = simulate(
          n = n, alpha = .x, tau = tau, params = params, B = BB,
          c = c, seed.offset = seed0 + .y, structure = "ind"
        )
        dplyr::mutate(out, alpha = .x, B = BB)
      })
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
  saveRDS(res_alpha, outfile)
}

# tidy
df = res_alpha %>%
  dplyr::filter(Method == "FDR_bound") %>%
  dplyr::mutate(
    FDP_mcse   = FDP_sd   / sqrt(B),
    Power_mcse = Power_sd / sqrt(B),
    B_f = factor(B, levels = Bs, labels = paste0("B = ", Bs))
  )
```

```{r B-plot}
# prep: keep only FDR_bound rows & add factor for B
df_B = res_alpha %>%
  filter(Method == "FDR_bound") %>%
  mutate(
    FDP_mcse   = FDP_sd   / sqrt(B),
    Power_mcse = Power_sd / sqrt(B),
    B_f = factor(
      B,
      levels = sort(unique(B)),
      labels = paste0("B = ", sort(unique(B)))
    )
  )

df_B_zoomed = subset(df_B, alpha <= 0.25)

# 3-color palette for the three B levels
B_levels = levels(df_B_zoomed$B_f)
base_cols = c("#0072B2", "#009E73", "#D55E00")  # blue / green / orange
cols_B = setNames(base_cols[seq_along(B_levels)], B_levels)

params_txt = paste0("(", paste(params, collapse = ", "), ")")

fdr_B = ggplot(df_B_zoomed, aes(x = alpha, y = FDP_mean, color = B_f, fill = B_f, group = B_f)) +
  geom_ribbon(
    aes(ymin = pmax(0, FDP_mean - FDP_mcse),
        ymax = pmin(1, FDP_mean + FDP_mcse)),
    alpha = 0.12, color = NA
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", linewidth = 1) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "B", values = cols_B) +
  scale_fill_manual(name = "B", values = cols_B) +
  labs(
    title = glue("Empirical FDR over alpha. n = {n}, parameters = {params_txt}, tau = {tau}, c = {c}"),
    x = expression(alpha),
    y = "Empirical FDR",
    color = "B", fill = "B",
    subtitle = glue("We used Iterations = ({paste(sort(unique(df_B$B)), collapse = ', ')}) for simulation.")
  ) +
  theme_linedraw(base_size = 13)

power_B = ggplot(df_B_zoomed, aes(x = alpha, y = Power_mean, color = B_f, fill = B_f, group = B_f)) +
  geom_ribbon(
    aes(ymin = pmax(0, Power_mean - Power_mcse),
        ymax = pmin(1, Power_mean + Power_mcse)),
    alpha = 0.12, color = NA
  ) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "B", values = cols_B) +
  scale_fill_manual(name = "B", values = cols_B) +
  labs(
    title = glue("Power vs α. n = {n}, parameters = {params_txt}, tau = {tau}, c = {c}"),
    x = expression(alpha),
    y = "Power",
    color = "B", fill = "B",
    subtitle = glue("We used Iterations = ({paste(sort(unique(df_B$B)), collapse = ', ')}) for simulation.")
  ) +
  theme_linedraw(base_size = 13)

fdr_B
power_B
```

# FDP, Power and Number of Discoveries

## FDP

```{r FDP}
B = 100 # default

params_txt = paste0("(", paste(params, collapse = ", "), ")")
outfile = "rds/power_ntar.rds"

# --- cache or compute ---
if (file.exists(outfile)) {
  res_alpha = readRDS(outfile)
} else {
  res_alpha = map_dfr(alphas, function(a) {
    out = simulate(n = n, alpha = a, tau = tau, params = params,
                   B = B, c = c, seed.offset = seed0, structure = "independent")
    out %>% mutate(alpha = a)
  })
  saveRDS(res_alpha, outfile)
}

# --- tidy for plotting ---
cols = c("FDR_bound" = "#009E73",  # green
         "Storey"    = "#E69F00",  # orange
         "BH"        = "#56B4E9",  # sky blue
         "BY"        = "#CC79A7")  # purple

df = res_alpha %>%
  mutate(
    Method     = factor(Method, levels = names(cols)),
    FDP_mcse   = FDP_sd   / sqrt(B),
    Power_mcse = Power_sd / sqrt(B)
  )

df = subset(df, alpha <= 0.25)

# --- FDP vs alpha plot (all methods) ---
p_fdp = ggplot(df, aes(x = alpha, y = FDP_mean, color = Method, fill = Method)) +
  geom_ribbon(aes(ymin = pmax(0, FDP_mean - FDP_mcse),
                  ymax = pmin(1, FDP_mean + FDP_mcse)),
              alpha = 0.12, color = NA) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = cols) +
  scale_fill_manual(values = cols) +
  labs(
    title = "Independent Structure",
    x = expression(alpha),
    y = "Empirical FDR",
    color = "Method", fill = "Method", subtitle = "Empirical FDR vs α"
  ) +
  theme_linedraw(base_size = 13) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed",
              linewidth = 0.7, color = "black", alpha = 0.7, show.legend = FALSE) +
  theme(legend.position = "none")

p_fdp
```

## Power

```{r power}
B = 100 # default

params_txt = paste0("(", paste(params, collapse = ", "), ")")
outfile = "rds/power_ntar.rds"

# --- cache or compute ---
if (file.exists(outfile)) {
  res_alpha = readRDS(outfile)
} else {
  res_alpha = map_dfr(alphas, function(a) {
    out = simulate(n = n, alpha = a, tau = tau, params = params,
                   B = B, c = c, seed.offset = seed0, structure = "independent")
    out %>% mutate(alpha = a)
  })
  saveRDS(res_alpha, outfile)
}

# --- tidy for plotting ---
cols = c("FDR_bound" = "#009E73",  # green
         "Storey"    = "#E69F00",  # orange
         "BH"        = "#56B4E9",  # sky blue
         "BY"        = "#CC79A7")  # purple

df = res_alpha %>%
  mutate(
    Method     = factor(Method, levels = names(cols)),
    FDP_mcse   = FDP_sd   / sqrt(B),
    Power_mcse = Power_sd / sqrt(B)
  )

df = subset(df, alpha <= 0.25)

# --- Power vs alpha plot (all methods) ---
p_power = ggplot(df, aes(x = alpha, y = Power_mean, color = Method, fill = Method)) +
  geom_ribbon(aes(ymin = pmax(0, Power_mean - Power_mcse),
                  ymax = pmin(1, Power_mean + Power_mcse)),
              alpha = 0.12, color = NA) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = cols) +
  scale_fill_manual(values = cols) +
  labs(
    title = "Independent Structure",
    x = expression(alpha),
    y = "Mean Power",
    color = "Method", fill = "Method", subtitle = "Power vs α"
  ) +
  theme_linedraw(base_size = 13)

p_power
```

## Number of Discoveries

```{r nT-plot}
# load cached results if not already present
if (!exists("res_alpha")) {
  res_alpha = readRDS("rds/power_ntar.rds")
}

# palette (same as before)
cols = c("FDR_bound" = "#009E73",
         "Storey"    = "#E69F00",
         "BH"        = "#56B4E9",
         "BY"        = "#CC79A7")

params_txt = paste0("(", paste(params, collapse = ", "), ")")

df = res_alpha %>%
  mutate(
    Method   = factor(Method, levels = names(cols)),
    R_mcse   = R_sd / sqrt(B)
  )

df = subset(df, alpha <= 0.25)

p_disc = ggplot(df, aes(x = alpha, y = R_mean, color = Method, fill = Method)) +
  geom_ribbon(aes(ymin = pmax(0, R_mean - R_mcse),
                  ymax = R_mean + R_mcse),
              alpha = 0.12, color = NA) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = cols) +
  scale_fill_manual(values = cols) +
  labs(
    title = glue("# Discoveries vs α. n = {n}, parameters = {params_txt}, tau = {tau}, c = {c}, M = {B}"),
    x = expression(alpha),
    y = "Mean # Discoveries",
    color = "Method", fill = "Method"
  ) +
  theme_linedraw(base_size = 13)

p_disc
```

# Variability of Empirical FDR and Power

## Against $\tau$

```{r}
res_alpha = readRDS("rds/tau_sim.rds")

df_tau = res_alpha %>%
  dplyr::filter(Method == "FDR_bound") %>%
  dplyr::mutate(
    # MCSEs for the MEANs (already what you had)
    FDP_mcse_mean   = FDP_sd   / sqrt(iters),
    Power_mcse_mean = Power_sd / sqrt(iters),
    # MCSEs for the SDs (delta-method / normal approx)
    FDP_sd_mcse   = FDP_sd   / sqrt(2 * pmax(iters - 1, 1)),
    Power_sd_mcse = Power_sd / sqrt(2 * pmax(iters - 1, 1))
  ) %>%
  dplyr::mutate(
    tau_f = factor(
      tau,
      levels = sort(unique(tau)),
      labels = paste0("τ = ", sort(unique(tau)))
    )
  )

df_tau = subset(df_tau, alpha <= 0.25)

# Color palette (expand/recycle if needed)
tau_levels = levels(df_tau$tau_f)
base_cols = c("#0072B2", "#009E73", "#D55E00")  # blue/green/orange
cols_tau = setNames(rep(base_cols, length.out = length(tau_levels)), tau_levels)

params_txt = paste0("(", paste(params, collapse = ", "), ")")

# --- FDR SD plot with MCSE ribbon ---
fdr_tau_var =
  ggplot(df_tau, aes(x = alpha, y = FDP_sd, color = tau_f)) +
  # Ribbon for SD ± MCSE(SD)
  geom_ribbon(
    aes(ymin = pmax(FDP_sd - FDP_sd_mcse, 0),
        ymax = FDP_sd + FDP_sd_mcse,
        fill = tau_f),
    alpha = 0.15, colour = NA
  ) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "τ", values = cols_tau) +
  scale_fill_manual(name = "τ", values = cols_tau) +
  labs(
    title = glue("FDP SD vs α. n = {n}, parameters = {params_txt}, c = {c}, M = {iters}"),
    subtitle = glue("τ values: {paste(sort(unique(df_tau$tau)), collapse = ', ')}"),
    x = expression(alpha),
    y = "FDP SD"
  ) +
  theme_linedraw(base_size = 13)

# --- Power SD plot with MCSE ribbon ---
power_tau_var =
  ggplot(df_tau, aes(x = alpha, y = Power_sd, color = tau_f)) +
  geom_ribbon(
    aes(ymin = pmax(Power_sd - Power_sd_mcse, 0),
        ymax = Power_sd + Power_sd_mcse,
        fill = tau_f),
    alpha = 0.15, colour = NA
  ) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "τ", values = cols_tau) +
  scale_fill_manual(name = "τ", values = cols_tau) +
  labs(
    title = glue("Power SD vs α. n = {n}, parameters = {params_txt}, c = {c}, M = {iters}"),
    subtitle = glue("τ values: {paste(sort(unique(df_tau$tau)), collapse = ', ')}"),
    x = expression(alpha),
    y = "Power SD"
  ) +
  theme_linedraw(base_size = 13)

fdr_tau_var
power_tau_var
```

## Against $c$

```{r}
res_alpha = readRDS("rds/c_sim.rds")

# Prep: keep FDR_bound, compute MCSEs for SDs
df_c = res_alpha %>%
  dplyr::filter(Method == "FDR_bound") %>%
  dplyr::mutate(
    # MCSEs for SDs (delta method)
    FDP_sd_mcse   = FDP_sd   / sqrt(2 * pmax(iters - 1, 1)),
    Power_sd_mcse = Power_sd / sqrt(2 * pmax(iters - 1, 1)),
    c_f = factor(
      c,
      levels = sort(unique(c)),
      labels = paste0("c = ", sort(unique(c)))
    )
  )

# Zoom to small alphas if desired
df_c_zoomed = subset(df_c, alpha <= 0.25)

# Palette
c_levels = levels(df_c_zoomed$c_f)
base_cols = c("#0072B2", "#009E73", "#D55E00", "slateblue")
cols_c = setNames(rep(base_cols, length.out = length(c_levels)), c_levels)

params_txt = paste0("(", paste(params, collapse = ", "), ")")

# --- FDP SD with MCSE(SD) ribbon ---
fdr_c_var = ggplot(df_c_zoomed, aes(x = alpha, y = FDP_sd, color = c_f)) +
  geom_ribbon(
    aes(ymin = pmax(FDP_sd - FDP_sd_mcse, 0),
        ymax = FDP_sd + FDP_sd_mcse,
        fill = c_f),
    alpha = 0.15, colour = NA
  ) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "c", values = cols_c) +
  scale_fill_manual(name = "c", values = cols_c) +
  labs(
    title = glue("FDP SD vs α. n = {n}, parameters = {params_txt}, τ = {tau}, M = {iters}"),
    subtitle = "c values: 0, 0.5, 1, 2",
    x = expression(alpha),
    y = "FDP SD",
    color = "c", fill = "c"
  ) +
  theme_linedraw(base_size = 13)

# --- Power SD with MCSE(SD) ribbon ---
power_c_var = ggplot(df_c_zoomed, aes(x = alpha, y = Power_sd, color = c_f)) +
  geom_ribbon(
    aes(ymin = pmax(Power_sd - Power_sd_mcse, 0),
        ymax = Power_sd + Power_sd_mcse,
        fill = c_f),
    alpha = 0.15, colour = NA
  ) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "c", values = cols_c) +
  scale_fill_manual(name = "c", values = cols_c) +
  labs(
    title = glue("Power SD vs α. n = {n}, parameters = {params_txt}, τ = {tau}, M = {iters}"),
    subtitle = "c values: 0, 0.5, 1, 2",
    x = expression(alpha),
    y = "Power SD",
    color = "c", fill = "c"
  ) +
  theme_linedraw(base_size = 13)

fdr_c_var
power_c_var
```

### Plot against FDR Bound

```{r}
# read c-sim results
res_c = readRDS("rds/c_sim.rds")

# keep only the FDR_bound *procedure* rows
df = res_c %>%
  filter(Method == "FDR_bound") %>%
  mutate(
    # harmonise naming in case your metrics used FDR_mean before renaming to FDR_bound_mean
    FDRb_mean = dplyr::coalesce(FDR_bound_mean, FDP_mean),
    FDRb_sd   = dplyr::coalesce(FDR_bound_sd, FDP_sd),  # if you have it; otherwise will be NA
    FDP_mcse_mean   = FDP_sd   / sqrt(iters),
    FDRb_mcse_mean  = FDRb_sd  / sqrt(iters),
    c_f = factor(c, levels = sort(unique(c)), labels = paste0("c = ", sort(unique(c))))
  )

# zoom (optional)
dfz = dplyr::filter(df, alpha <= 0.25)

# palette
c_levels = levels(dfz$c_f)
base_cols = c("#0072B2", "#009E73", "#D55E00", "slateblue")
cols_c = setNames(rep(base_cols, length.out = length(c_levels)), c_levels)

params_txt = paste0("(", paste(params, collapse = ", "), ")")

# --- Panel A: FDR_bound (solid) vs FDP (dashed), with MCSE ribbons for means ---
p_c_bound_vs_fdp =
  ggplot(dfz, aes(x = alpha, colour = c_f, fill = c_f, group = c_f)) +
  # FDP mean ribbon
  geom_ribbon(aes(ymin = pmax(0, FDP_mean - FDP_mcse_mean),
                  ymax = pmin(1, FDP_mean + FDP_mcse_mean)),
              alpha = 0.10, colour = NA) +
  # FDR_bound mean ribbon (only if sd exists)
  geom_ribbon(aes(ymin = pmax(0, FDRb_mean - FDRb_mcse_mean),
                  ymax = pmin(1, FDRb_mean + FDRb_mcse_mean)),
              alpha = 0.10, colour = NA, inherit.aes = TRUE, show.legend = FALSE,
              data = ~subset(., !is.na(FDRb_mcse_mean))) +
  # lines
  geom_line(aes(y = FDRb_mean), linewidth = 1) +
  geom_line(aes(y = FDP_mean), linewidth = 1, linetype = "dashed") +
  scale_color_manual(values = cols_c, name = "c") +
  scale_fill_manual(values = cols_c,  name = "c") +
  labs(
    title = glue("FDR_bound vs FDP over α. n = {n}, params = {params_txt}, τ = {tau}, M = {iters}"),
    subtitle = "Zoom: α ≤ 0.25 • Ribbons show MCSE of the mean",
    x = expression(alpha), y = "Rate"
  ) +
  theme_linedraw(base_size = 13) +
  facet_wrap(~ c_f)

# --- Panel B: Gap = mean(FDR_bound) − mean(FDP) (want ≥ 0 on average) ---
df_gap = dfz %>%
  mutate(
    gap = FDRb_mean - FDP_mean,
    gap_mcse = sqrt(FDRb_mcse_mean^2 + FDP_mcse_mean^2)  # independent-reps approximation
  )

p_c_gap =
  ggplot(df_gap, aes(x = alpha, y = gap, colour = c_f, fill = c_f)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_ribbon(aes(ymin = gap - gap_mcse, ymax = gap + gap_mcse), alpha = 0.12, colour = NA) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = cols_c, name = "c") +
  scale_fill_manual(values = cols_c,  name = "c") +
  labs(
    title = glue("Gap: FDR_bound − FDP. n = {n}, params = {params_txt}, τ = {tau}, M = {iters}"),
    subtitle = "Positive values support that the bound exceeds FDP on average",
    x = expression(alpha), y = "Gap"
  ) +
  theme_linedraw(base_size = 13) +
  facet_wrap(~ c_f)

p_c_bound_vs_fdp
p_c_gap
```

It seems that average FDR bound does exceed average FDP. The procedure is not overly conservative: the bound closely tracks the empirical FDP. The gap is small but positive, and grows slightly with $\alpha$. This supports that the FDR bound method achieves its guarantee in expectation, even if FDR control is not enforced replicate by replicate.

We can check using higher simulation count to see if there are still dips.

```{r}
cs = c(0, 0.5, 1, 2)
iters = 1000
seed0 = 123

outfile = "rds/c_sim2.rds"
future::plan(multisession, workers = max(1, parallel::detectCores() - 1))

if (file.exists(outfile)) {
  res_alpha = readRDS(outfile)
} else {
  res_alpha = furrr::future_map2_dfr(
    alphas, seq_along(alphas),
    ~{
      purrr::map_dfr(cs, function(cc) {
        out = simulate(
          n, alpha = .x, tau = tau, params = params, B = iters,
          c = cc, seed.offset = seed0 + .y, structure = "ind"
        )
        dplyr::mutate(out, alpha = .x, c = cc)
      })
    },
    .options = furrr::furrr_options(seed = TRUE)
  )
  saveRDS(res_alpha, outfile)
}
```

```{r}
# tidy
# read c-sim results
res_c = readRDS("rds/c_sim2.rds")

# keep only the FDR_bound *procedure* rows
df = res_c %>%
  filter(Method == "FDR_bound") %>%
  mutate(
    # harmonise naming in case your metrics used FDR_mean before renaming to FDR_bound_mean
    FDRb_mean = dplyr::coalesce(FDR_bound_mean, FDP_mean),
    FDRb_sd   = dplyr::coalesce(FDR_bound_sd, FDP_sd),  # if you have it; otherwise will be NA
    FDP_mcse_mean   = FDP_sd   / sqrt(iters),
    FDRb_mcse_mean  = FDRb_sd  / sqrt(iters),
    c_f = factor(c, levels = sort(unique(c)), labels = paste0("c = ", sort(unique(c))))
  )

# zoom (optional)
dfz = dplyr::filter(df, alpha <= 0.25)

# palette
c_levels = levels(dfz$c_f)
base_cols = c("#0072B2", "#009E73", "#D55E00", "slateblue")
cols_c = setNames(rep(base_cols, length.out = length(c_levels)), c_levels)

params_txt = paste0("(", paste(params, collapse = ", "), ")")

# --- Panel A: FDR_bound (solid) vs FDP (dashed), with MCSE ribbons for means ---
p_c_bound_vs_fdp =
  ggplot(dfz, aes(x = alpha, colour = c_f, fill = c_f, group = c_f)) +
  # FDP mean ribbon
  geom_ribbon(aes(ymin = pmax(0, FDP_mean - FDP_mcse_mean),
                  ymax = pmin(1, FDP_mean + FDP_mcse_mean)),
              alpha = 0.10, colour = NA) +
  # FDR_bound mean ribbon (only if sd exists)
  geom_ribbon(aes(ymin = pmax(0, FDRb_mean - FDRb_mcse_mean),
                  ymax = pmin(1, FDRb_mean + FDRb_mcse_mean)),
              alpha = 0.10, colour = NA, inherit.aes = TRUE, show.legend = FALSE,
              data = ~subset(., !is.na(FDRb_mcse_mean))) +
  # lines
  geom_line(aes(y = FDRb_mean), linewidth = 1) +
  geom_line(aes(y = FDP_mean), linewidth = 1, linetype = "dashed") +
  scale_color_manual(values = cols_c, name = "c") +
  scale_fill_manual(values = cols_c,  name = "c") +
  labs(
    title = glue("FDR_bound vs FDP over α. n = {n}, params = {params_txt}, τ = {tau}, M = {iters}"),
    subtitle = "Zoom: α ≤ 0.25 • Ribbons show MCSE of the mean",
    x = expression(alpha), y = "Rate"
  ) +
  theme_linedraw(base_size = 13) +
  facet_wrap(~ c_f)

# --- Panel B: Gap = mean(FDR_bound) − mean(FDP) (want ≥ 0 on average) ---
df_gap = dfz %>%
  mutate(
    gap = FDRb_mean - FDP_mean,
    gap_mcse = sqrt(FDRb_mcse_mean^2 + FDP_mcse_mean^2)  # independent-reps approximation
  )

p_c_gap =
  ggplot(df_gap, aes(x = alpha, y = gap, colour = c_f, fill = c_f)) +
  geom_hline(yintercept = 0, linetype = "dotted") +
  geom_ribbon(aes(ymin = gap - gap_mcse, ymax = gap + gap_mcse), alpha = 0.12, colour = NA) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = cols_c, name = "c") +
  scale_fill_manual(values = cols_c,  name = "c") +
  labs(
    title = glue("Gap: FDR_bound − FDP. n = {n}, params = {params_txt}, τ = {tau}, M = {iters}"),
    subtitle = "Positive values support that the bound exceeds FDP on average",
    x = expression(alpha), y = "Gap"
  ) +
  theme_linedraw(base_size = 13) +
  facet_wrap(~ c_f)

p_c_bound_vs_fdp
p_c_gap


```

## Against $n$

```{r}
res_alpha = readRDS("rds/n_sim.rds")

df_n = res_alpha %>%
  filter(Method == "FDR_bound") %>%
  mutate(
    # MCSEs for SDs (delta method)
    FDP_sd_mcse   = FDP_sd   / sqrt(2 * pmax(iters - 1, 1)),
    Power_sd_mcse = Power_sd / sqrt(2 * pmax(iters - 1, 1)),
    n_f = factor(
      n,
      levels = sort(unique(n)),
      labels = paste0("n = ", sort(unique(n)))
    )
  )

# zoom (note: alpha is lowercase)
df_n_zoomed = subset(df_n, alpha <= 0.25)

# palette
n_levels = levels(df_n_zoomed$n_f)
base_cols = c("#0072B2", "#009E73", "#D55E00")
cols_n = setNames(rep(base_cols, length.out = length(n_levels)), n_levels)

params_txt = paste0("(", paste(params, collapse = ", "), ")")

# --- FDP SD with MCSE(SD) ribbon ---
fdr_n_var = ggplot(df_n_zoomed, aes(x = alpha, y = FDP_sd, color = n_f)) +
  geom_ribbon(
    aes(ymin = pmax(FDP_sd - FDP_sd_mcse, 0),
        ymax = FDP_sd + FDP_sd_mcse,
        fill = n_f),
    alpha = 0.15, colour = NA
  ) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "n", values = cols_n) +
  scale_fill_manual(name = "n", values = cols_n) +
  labs(
    title = glue("FDP SD vs α. parameters = {params_txt}, τ = {tau}, c = {c}, M = {iters}"),
    subtitle = glue("n values: {paste(sort(unique(df_n$n)), collapse = ', ')} (zoomed to α ≤ 0.25)"),
    x = expression(alpha),
    y = "FDP SD",
    color = "n", fill = "n"
  ) +
  theme_linedraw(base_size = 13)

# --- Power SD with MCSE(SD) ribbon ---
power_n_var = ggplot(df_n_zoomed, aes(x = alpha, y = Power_sd, color = n_f)) +
  geom_ribbon(
    aes(ymin = pmax(Power_sd - Power_sd_mcse, 0),
        ymax = Power_sd + Power_sd_mcse,
        fill = n_f),
    alpha = 0.15, colour = NA
  ) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "n", values = cols_n) +
  scale_fill_manual(name = "n", values = cols_n) +
  labs(
    title = glue("Power SD vs α. parameters = {params_txt}, τ = {tau}, c = {c}, M = {iters}"),
    subtitle = glue("n values: {paste(sort(unique(df_n$n)), collapse = ', ')} (zoomed to α ≤ 0.25)"),
    x = expression(alpha),
    y = "Power SD",
    color = "n", fill = "n"
  ) +
  theme_linedraw(base_size = 13)

fdr_n_var
power_n_var
```

## Against $B$

```{r}
res_alpha = readRDS("rds/B_sim.rds")

# Keep FDR_bound, compute MCSEs for SDs (delta method), and factor for B
df_B = res_alpha %>%
  filter(Method == "FDR_bound") %>%
  mutate(
    FDP_sd_mcse   = FDP_sd   / sqrt(2 * pmax(B - 1, 1)),
    Power_sd_mcse = Power_sd / sqrt(2 * pmax(B - 1, 1)),
    B_f = factor(
      B,
      levels = sort(unique(B)),
      labels = paste0("B = ", sort(unique(B)))
    )
  )

# zoom to α ≤ 0.25 (use df_B for full range)
df_B_zoomed = subset(df_B, alpha <= 0.25)

# palette
B_levels = levels(df_B_zoomed$B_f)
base_cols = c("#0072B2", "#009E73", "#D55E00")
cols_B = setNames(rep(base_cols, length.out = length(B_levels)), B_levels)

params_txt = paste0("(", paste(params, collapse = ", "), ")")

# --- FDP SD with MCSE(SD) ribbon ---
fdr_B_var = ggplot(df_B_zoomed, aes(x = alpha, y = FDP_sd, color = B_f)) +
  geom_ribbon(
    aes(ymin = pmax(FDP_sd - FDP_sd_mcse, 0),
        ymax = FDP_sd + FDP_sd_mcse,
        fill = B_f),
    alpha = 0.15, colour = NA
  ) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "B", values = cols_B) +
  scale_fill_manual(name = "B", values = cols_B) +
  labs(
    title = glue("FDP SD vs α. n = {n}, parameters = {params_txt}, τ = {tau}, c = {c}"),
    subtitle = glue("M = {paste(sort(unique(df_B$B)), collapse = ', ')}"),
    x = expression(alpha),
    y = "FDP SD",
    color = "B", fill = "B"
  ) +
  theme_linedraw(base_size = 13)

# --- Power SD with MCSE(SD) ribbon ---
power_B_var = ggplot(df_B_zoomed, aes(x = alpha, y = Power_sd, color = B_f)) +
  geom_ribbon(
    aes(ymin = pmax(Power_sd - Power_sd_mcse, 0),
        ymax = Power_sd + Power_sd_mcse,
        fill = B_f),
    alpha = 0.15, colour = NA
  ) +
  geom_line(linewidth = 1) +
  scale_color_manual(name = "B", values = cols_B) +
  scale_fill_manual(name = "B", values = cols_B) +
  labs(
    title = glue("Power SD vs α. n = {n}, parameters = {params_txt}, τ = {tau}, c = {c}"),
    subtitle = glue("M = {paste(sort(unique(df_B$B)), collapse = ', ')}"),
    x = expression(alpha),
    y = "Power SD",
    color = "B", fill = "B"
  ) +
  theme_linedraw(base_size = 13)

fdr_B_var
power_B_var
```